ANTs-Flavored Tract-Based Spatial Statistics

### 2011-03-25 ANTsTBSS/paper.tex
Minor edits, new figure.
Including Phil's suggestions.

\begin{enumerate}
  \item Build the anatomical template from  the population sample or sample subset using SyGN.
  \item If only using a sample subset for template creation, warp the remaining
        anatomical data to the template and save the resulting transformation for each         
        subject indexed by $i$ ($T^i_{template}$).
  \item Find the optimal transformation from each
        subject's FA to the anatomical template:
  \begin{enumerate}
  \item For each anatomical image, segment the white matter (in ANTs one can use N4 bias correction and Atropos $n$-tissue segmentation).
  \item Rigidly transform each subject's average DWI to the bias-corrected anatomical
        image ($R^i_{dwi}$).
  \item Perform distortion correction on the FA to anatomical mapping 
        by localized nonrigid registration
        of rigidly transformed FA map ($=R^i_{dwi} \mathrm{FA}$) using
        the white matter mask obtained in step 3 ($T^i_{FA}$).
  \item For each subject, concatenate the transforms which maps each
        subject's FA to the template ($T^i_{total} = T^i_{template} \circ 
        T^i_{FA} \circ R^i_{dwi} \mathrm{FA}$).
  \item Warp each subject's FA or MD map to the template using $T^i_{total}$.
  \end{enumerate}      
  \item Perform the remaining portions of the standard TBSS pipeline:
    \begin{enumerate}
      \item Create the mean FA image from all template-aligned FA images.
      \item Produce the white matter skeleton from the mean FA image.
      \item Create skeleton distance map and project all FA data onto the skeleton.
      \item Run the statistical analysis.
    \end{enumerate}
\end{enumerate}

@@

The tensorial data associated with each of the following data sets were reconstructed from the diffusion weighted sequence using Camino%
\footnote{
http://web4.cs.ucl.ac.uk/research/medic/camino/pmwiki/
}---an open source toolkit for diffusion MRI processing and analysis \cite{Cook2006} in combination with ANTs-based registration for motion correction of the DTI sequence.

@@

A third data set was obtained as port of a larger effort investigating the relationship between various neuroimaging indices and cognitive and functional abilities in chronic survivors of TBI (principal investigator: John Whyte). A more detailed data description can be found in \cite{Avants2008}. The following is a brief description relevant to our particular study.  
17 controls and 16 patients with TBI were used for the analysis presented in this
work.  Each patient had a history of non-penetrating traumatic brain injury 
of at least moderate severity defined by significant and well-documented
loss or alteration of consciousness following injury in addition to meeting
several other exclusionary criteria.  The healthy volunteers were matched in terms of age, gender, ethnicity, handedness, and
years of education.
High resolution T1-weighted anatomic images were obtained using 3D MP-RAGE 
imaging sequence using the following acquisition parameters: TR = 1620 ms, 
TI = 950 ms, TE = 3 ms, flip angle = 15$^\circ$, 160 contiguous slices of 
1.0 mm thickness, FOV = 192 $\times$ 256 mm$^2$, matrix = 192 $\times$ 256, 
1NEX with a scan time of 6 minutes, and voxel size = 1 mm$^3$.  30-directional
DTI images were also obtained

@@

%Comparison of areas of significant difference between the two groups for both methodologies can be seen on several corresponding axial slices in Figure \ref{fig:hoonData}.  The majority of the areas of significance overlap between the two analyses although the yellow arrow points to one difference situated in the right frontal lobe.

>>>>

\begin{enumerate}
  \item Build the anatomical template from  the population sample or sample subset using SyGN.
  \item If only using a sample subset for template creation, warp the remaining
        anatomical data to the template and save the resulting transformation for each         
        subject indexed by $i$ ($T^i_{template}$).
  \item Find the optimal transformation from each
        subject's FA to the anatomical template:
  \begin{enumerate}
  \item For each anatomical image, segment the white matter (in ANTs one can use N4 bias correction and Atropos $n$-tissue segmentation).
  \item Rigidly transform each subject's average DWI to the bias-corrected anatomical
        image ($R^i_{dwi}$).
  \item Perform distortion correction on the FA to anatomical mapping 
        by localized nonrigid registration
        of rigidly transformed FA map ($=R^i_{dwi} \mathrm{FA}$) using
        the white matter mask obtained in step 3(a) ($T^i_{FA}$).
  \item For each subject, concatenate the transforms which maps each
        subject's FA to the template ($T^i_{total} = T^i_{template} \circ 
        T^i_{FA} \circ R^i_{dwi} \mathrm{FA}$).
  \item Warp each subject's FA or MD map to the template using $T^i_{total}$.
  \end{enumerate}      
  \item Perform the remaining portions of the standard TBSS pipeline:
    \begin{enumerate}
      \item Create the mean FA image from all template-aligned FA images.
      \item Produce the white matter skeleton from the mean FA image.
      \item Create skeleton distance map and project all FA data onto the skeleton.
      \item Run the statistical analysis.
    \end{enumerate}
\end{enumerate}

@@

The tensorial data associated with each of the following data sets were reconstructed from the diffusion weighted sequence using Camino%
\footnote{
http://camino.org.uk/
}---an open source toolkit for diffusion MRI processing and analysis \cite{Cook2006} in combination with ANTs-based registration for motion correction of the DTI sequence.

@@

A third data set was obtained as port of a larger effort investigating the relationship between various neuroimaging indices and cognitive and functional abilities in long-term survivors of TBI (principal investigator: John Whyte). A more detailed data description can be found in \cite{Avants2008}. The following is a brief description relevant to our particular study.  
17 controls and 16 patients with TBI were used for the analysis presented in this
work.  Each patient had a history of non-penetrating traumatic brain injury 
of at least moderate severity defined by significant and well-documented
loss or alteration of consciousness following injury in addition to meeting
several other exclusionary criteria.  The healthy volunteers were matched in terms of age, gender, ethnicity, handedness, and
years of education.
High resolution T1-weighted anatomic images were obtained using 3D MP-RAGE 
imaging sequence using the following acquisition parameters: TR = 1620 ms, 
TI = 950 ms, TE = 3 ms, flip angle = 15$^\circ$, 160 contiguous slices of 
1.0 mm thickness, FOV = 192 $\times$ 256 mm$^2$, matrix = 192 $\times$ 256, 
1NEX with a scan time of 6 minutes, and voxel size = 1 mm$^3$.  30-directional
DTI images were also obtained

@@

%
%Comparison of areas of significant difference between the two groups for both methodologies can be seen on several corresponding axial slices in Figure \ref{fig:hoonData}.  The majority of the areas of significance overlap between the two analyses although the yellow arrow points to one difference situated in the right frontal lobe.
